---
title: "Activities"
output: html_document
---
1. Code for reading in the dataset and/or processing the data
```{r}
data <- read.csv("/Users/javis/Downloads/repdata_data_activity/activity.csv")
head(data)
summary(data)
str(data)
```
2. Histogram of the total number of steps taken each day

```{r, echo = TRUE}
library(ggplot2)
totalsteps <- with(data, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(totalsteps) <- c("date", "steps")
head(totalsteps)
hist(totalsteps$steps, main = "Total steps taken per day", xlab = "Steps taken per day", col = "red")
```
3. Mean and median number of steps taken each day

```{r pressure, echo=FALSE}
median(totalsteps$steps)
mean(totalsteps$steps)

```
4. Time series plot of the average number of steps taken
```{r, echo = TRUE}
average_daily_activity <- aggregate(data$steps, by=list(data$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="darkblue", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```
5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval

```

6. Code to describe and show a strategy for imputing missing data
```{r}
imputed_steps <- average_daily_activity$mean[match(data$interval, average_daily_activity$interval)]
activity_imputed <- transform(data, steps = ifelse(is.na(data$steps), yes = imputed_steps, no = data$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")

```
7. Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo = TRUE}
hist(total_steps_imputed$daily_steps, main = "Total steps taken per day", xlab = "Steps taken per day", col = "red")
```
Also, for compare this transform we must do:
```{r}
mean(total_steps_imputed$daily_steps)
median(total_steps_imputed$daily_steps)
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
First, we must to transform our data in order to get the plot that we need:
```{r}
data$date <- as.Date(strptime(data$date, format="%Y-%m-%d"))
data$datetype <- sapply(data$date, function(x) {
  if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
  {y <- "Weekend"} else 
  {y <- "Weekday"}
  y
})
activity_by_date <- aggregate(steps~interval + datetype, data, mean, na.rm = TRUE)
```
Now, the desired plot looks like:
```{r, echo = TRUE}
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
  geom_line() +
  labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
  facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```
